
---
title: "Coolest Thing You've Learned So Far"
author: "Daniel Craig"
date: '2022-06-08'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(rmarkdown)
knitr::opts_chunk$set(fig.path = "../images/")
```
## Coolest Thing I've Learned in R

So I think the coolest thing I've learned in R is honestly extremely simple.

The nature in which the language is interpreted is so simple it blows my mind. Particularly when I am rebuilding/morphing/manipulating data sets its just easy.

For example, when we were doing our caretPackage and model training homework, I wanted to remove the date column from my dataset. I'm sure there's a lot of ways to pop off one column, but the below lines blew me away:

`seoulBike <- seoulBike[ ,!(names(seoulBike) %in% c("Date"))]`

Any column with the name Date in it? Gone! All in the same line as the one I set the object again. I come from an SQL background and the ability to manipulate objects is a bit more crude so this is just very nice for me. I'm sure dev's from other languages don't think this is a big deal but it really just makes me appreciate how eloquently some coding languages R.

Just to ensure that you don't think I picked this because it was easy, I have gone ahead and attempted a plot below.


```{r , eval=TRUE}
seoulBike <- read_csv(file = "../_Rmd/_datasets/SeoulBikeData.csv", locale=locale(encoding="latin1"))
seoulBike <- seoulBike[ ,!(names(seoulBike) %in% c("Date"))]

names(seoulBike) <- c('RentBikeCount','Hour','Temp','Humid','WindSpeed','Visi','DewPointTemp','SolarRad','Rainfall','Snowfall','Seasons','Holiday','WorkDay','HighRentDay')

names(seoulBike)

correlation <- cor(seoulBike$RentBikeCount, seoulBike$Temp)

g <- ggplot(data = seoulBike, aes(x = RentBikeCount, y = Temp)) 
g + geom_point() + geom_smooth() + geom_smooth(method = glm, col = "Red") +
  geom_text(x=2500, y = 5, size = 5, label = paste0("Correlation = ", round(correlation, 2)))
```






rmarkdown::render("_Rmd/2022-06-30-CoolThing.Rmd",output_format = md_document("markdown_github"),output_dir = "_posts", output_options = list(keep_html=FALSE))


